---
title: "Regional Concentration (Canada)"
output: html_document
---
```{r include=FALSE}
source("data_prep.R")
library(dplyr)
library(plotly)
library(ggplot2)
library(scales)
```

```{r interactive-top30-fixed, echo=FALSE, fig.cap="Figure 4: Regional concentration of border-crossing volume at the top-30 busiest ports", echo=FALSE}
library(dplyr)
library(stringr)
library(leaflet)
library(tibble)

top30 <- df_1 %>%
  mutate(port_code = str_extract(`Port of Entry`, "^[0-9]{4}")) %>%
  group_by(port_code, `Port of Entry`) %>%
  summarise(Volume = sum(`Sum of Volume`, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(Volume)) %>%
  slice_head(n = 30) %>%
  rename(port_name = `Port of Entry`)

coords30 <- tribble(
  ~port_code, ~lon,       ~lat,
  "4971",     -79.6248,   43.6777,  # Pearson Intl T1
  "8212",     -73.5673,   45.5017,  # VIA – Traffic (Montreal)
  "4974",     -79.6248,   43.6777,  # Pearson Intl T3
  "3961",     -73.7408,   45.4706,  # Trudeau Intl Airport
  "8135",    -122.7579,   49.0018,  # Douglas Passenger (Peace Arch)
  "4101",     -78.9019,   42.9030,  # Fort Erie
  "4531",     -83.0364,   42.3140,  # Ambassador Bridge
  "8132",    -122.7579,   49.0018,  # Pacific Highway
  "4272",     -79.0742,   43.0896,  # Rainbow Bridge
  "4521",     -83.0307,   42.3090,  # Windsor Tunnel
  "4273",     -79.0660,   43.1214,  # Queenston Bridge
  "4401",     -82.4050,   42.9991,  # Blue Water Bridge
  "7011",    -114.0194,   51.1139,  # Calgary Intl Airport
  "3513",     -73.3440,   45.0080,  # Lacolle Route 15
  "4530",     -83.0364,   42.3140,  # Ambassador Bridge (Comm)
  "8173",     -73.3660,   45.0100,  # Huntingdon Traffic
  "4090",     -74.7306,   45.0014,  # Cornwall Traffic
  "8150",    -122.8270,   49.0980,  # Boundary Bay
  "4560",     -75.9310,   44.2690,  # Lansdowne Traffic
  "4410",     -84.3469,   46.5020,  # Sault Ste. Marie
  "8174",    -122.4840,   49.0367,  # Aldergrove Traffic
  "3281",     -73.3670,   45.0220,  # St.-Armand Traffic
  "8111",    -123.4260,   48.6450,  # Victoria Airport
  "4403",     -82.4050,   42.9991,  # Blue Water Bridge (Comm)
  "7023",    -113.5800,   53.3097,  # Edmonton Intl Airport
  "4780",     -93.4130,   48.6010,  # Fort Frances
  "2110",     -67.2710,   45.2040,  # St. Stephen Traffic
  "4312",     -75.6690,   45.3220,  # Ottawa Intl Airport
  "3141",     -72.0914,   45.0063,  # Stanstead Route 55
  "5021",     -97.2241,   49.0078   # Emerson West Lynne
)


df_plot <- left_join(top30, coords30, by = "port_code")

pal <- colorNumeric("Reds", domain = df_plot$Volume)

leaflet(df_plot) %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng = -95, lat = 55, zoom = 4) %>%
  addCircleMarkers(
    ~lon, ~lat,
    radius      = ~Volume / 800000,
    weight      = 1,
    color       = "white",
    fillColor   = ~pal(Volume),
    fillOpacity = 0.8,
    label       = ~paste0(port_name, " (", port_code, "): ", scales::comma(Volume)),
    labelOptions= labelOptions(direction = "auto")
  ) %>%
  addLegend(
    "bottomright",
    pal      = pal,
    values   = ~Volume,
    title    = "Total Crossings",
    labFormat= labelFormat(big.mark = ",")
  )
```

# Summary:
The map titled "Regional Concentration (Canada)" shows the total border-crossing volumes at the top 30 busiest Canadian ports of entry. Each circle represents a port, and its size and color intensity indicate the total number of crossings. The data reveals a strong concentration of crossings in southern Ontario, particularly near the U.S. border around the Great Lakes, with far fewer high-volume ports in western and eastern Canada. The busiest port is Pearson International Airport (Toronto) with over 18 million crossings, followed by Trudeau International Airport (Montreal) with over 13 million. So, it also aligns with the result shown in Figure 3 that air will be the second highest crossing volumes in Canada.