---
title: "Border Crossings by Region and Mode of Transportation"
output: html_document
---
```{r echo=FALSE}
source("data_prep.R")
library(dplyr)
library(plotly)
library(ggplot2)
library(scales)
```

```{r echo=FALSE}
library(dplyr)
library(plotly)
library(scales)  # for comma()

# Recompute totals by region and mode
df_regmode <- df_1 %>%
  group_by(Region, Mode) %>%
  summarise(Total = sum(`Sum of Volume`, na.rm = TRUE), .groups="drop")

# Plotly grouped bar with a log-y axis
plot_ly(
  df_regmode,
  x         = ~Region,
  y         = ~Total,
  color     = ~Mode,
  colors    = "Set2",
  type      = "bar",
  text      = ~comma(Total),
  hoverinfo = "text+legend"
) %>%
  layout(
    barmode = "group",
    title   = "Border Crossings by Region and Transportation Mode (Log Scale)",
    xaxis   = list(tickangle = -45),
    yaxis   = list(
                title = "Total Crossings (log scale) by Daily",
                type  = "log"
              ),
    legend = list(title = list(text = "<b>Mode</b>"))
  )
```
**Figure 2.** _Trend of Border Crossings by Region._ 

# Summary:
The bar chart illustrates total border crossings by region, showing Southern Ontario and the Pacific Region as the two most active regions and each has over 30 million crossings. The Greater Toronto Area follows. In contrast, the Quebec, Prairie, Northern Ontario, and Atlantic regions experience progressively fewer crossings, with the Atlantic Region having the lowest total volume. The dominance of Southern Ontario and the Pacific Region suggests that major urban centers and key border points significantly influence travel volumes.

```{r echo=FALSE}
library(dplyr)
library(stringr)
library(leaflet)
library(tibble)

# 1. Summarise top-30 port codes and volumes
top30 <- df_1 %>%
  mutate(port_code = str_extract(`Port of Entry`, "^[0-9]{4}")) %>%
  group_by(port_code) %>%
  summarise(Volume = sum(`Sum of Volume`, na.rm = TRUE), .groups="drop") %>%
  arrange(desc(Volume)) %>%
  slice_head(n = 30)

# 2. Inline coord table for those 30 codes
coords30 <- tribble(
  ~port_code, ~lon,       ~lat,
  "4971",     -79.6248,   43.6777,  # Pearson Intl T1
  "8212",     -73.5673,   45.5017,  # VIA – Traffic (Montreal)
  "4974",     -79.6248,   43.6777,  # Pearson Intl T3
  "3961",     -73.7408,   45.4706,  # Trudeau Intl Airport
  "8135",    -122.7579,   49.0018,  # Douglas Passenger (Peace Arch)
  "4101",     -78.9019,   42.9030,  # Fort Erie
  "4531",     -83.0364,   42.3140,  # Ambassador Bridge
  "8132",    -122.7579,   49.0018,  # Pacific Highway
  "4272",     -79.0742,   43.0896,  # Rainbow Bridge
  "4521",     -83.0307,   42.3090,  # Windsor Tunnel
  "4273",     -79.0660,   43.1214,  # Queenston Bridge
  "4401",     -82.4050,   42.9991,  # Blue Water Bridge
  "7011",    -114.0194,   51.1139,  # Calgary Intl Airport
  "3513",     -73.3440,   45.0080,  # Lacolle Route 15
  "4530",     -83.0364,   42.3140,  # Ambassador Bridge (Comm)
  "8173",     -73.3660,   45.0100,  # Huntingdon Traffic
  "4090",     -74.7306,   45.0014,  # Cornwall Traffic
  "8150",    -122.8270,   49.0980,  # Boundary Bay
  "4560",     -75.9310,   44.2690,  # Lansdowne Traffic
  "4410",     -84.3469,   46.5020,  # Sault Ste. Marie
  "8174",    -122.4840,   49.0367,  # Aldergrove Traffic
  "3281",     -73.3670,   45.0220,  # St.-Armand Traffic
  "8111",    -123.4260,   48.6450,  # Victoria Airport
  "4403",     -82.4050,   42.9991,  # Blue Water Bridge (Comm)
  "7023",    -113.5800,   53.3097,  # Edmonton Intl Airport
  "4780",     -93.4130,   48.6010,  # Fort Frances
  "2110",     -67.2710,   45.2040,  # St. Stephen Traffic
  "4312",     -75.6690,   45.3220,  # Ottawa Intl Airport
  "3141",     -72.0914,   45.0063,  # Stanstead Route 55
  "5021",     -97.2241,   49.0078   # Emerson West Lynne
)

# 3. Combine volumes & coords
df_plot <- left_join(top30, coords30, by = "port_code")

# 4. Warn if any missing
missing <- filter(df_plot, is.na(lon))
if(nrow(missing)) {
  warning("Missing coords for codes: ", paste(missing$port_code, collapse=", "))
}

# 5. Color palette
pal <- colorNumeric("Reds", domain = df_plot$Volume)

# 6. Build interactive map
leaflet(df_plot) %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng = -95, lat = 55, zoom = 4) %>%
  addCircleMarkers(
    ~lon, ~lat,
    radius      = ~Volume / 800000,
    weight      = 1,
    color       = "white",
    fillColor   = ~pal(Volume),
    fillOpacity = 0.8,
    label       = ~paste0(port_code, ": ", scales::comma(Volume)),
    labelOptions= labelOptions(direction = "auto")
  ) %>%
  addLegend(
    "bottomright",
    pal      = pal,
    values   = ~Volume,
    title    = "Total Crossings",
    labFormat= labelFormat(big.mark = ",")
  )
```
**Figure 2.** _Map of Top 30 Busiest Port Crossings._ 



